Multi-Resolution Neural Field Estimation

### 2011-11-04 Multiresolution_framework/ltx/manuscript.tex
commit for fist merge

%    \documentclass[journal,a4paper]{IEEEtran}
%  \documentclass[journal]{IEEEtran}

%      \documentclass[10pt,twocolumn,twoside]{IEEEtran}
     \documentclass[11pt,draftcls,onecolumn,peerreview]{IEEEtran} 
%         \topmargin       -6.0mm
%          \oddsidemargin      0mm
%          \evensidemargin     0mm
%          \textheight     223.5mm
%         \textwidth      170.0mm

@@

\usepackage{stfloats}
\usepackage{float}
\floatstyle{ruled}
%This bit is how to use hyperlink without stupid coloured rectangles around it
\usepackage{float}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor=black,filecolor=black,urlcolor=black,citecolor=black}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\newfloat{algorithm}{htbp}{loa}%[chapter]
\floatname{algorithm}{Algorithm}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model}

@@

\begin{abstract}
%\boldmath
We develop a multi-resolution approximation (MRA) framework for the integro-difference equation (IDE) neural field model based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract}
% Note that keywords are not normally used for peerreview papers.
 \begin{IEEEkeywords}
  Neural field model, multi-resolution approximation (MRA), Expectation Maximization (EM) algorithm, wavelets.
  \end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
%   \begin{center} \bfseries EDICS Category: SSP-IDEN \end{center}
%   \fi
%
%  For peerreview papers, this IEEEtran command inserts a page break and
%creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

\section{IDE Neural Field Model}
The stochastic IDE form of the Amari neural field  formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The term $e_t(\mathbf r)$ is a zero mean Gaussian disturbance, spatially coloured but temporally white, with covarianve function 
\begin{equation}
cov\left(e_{t}\left(\mathbf{r}\right),e_{t+t'}\left(\mathbf{r'}\right)\right)=\eta(\mathbf{r}-\mathbf{r'})\delta(t-t'),
\label{eq:FieldDisturbance}
\end{equation}
where $\delta(\cdot)$ is the Dirac-delta function. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$ \cite{VanRotterdam1982,Murphy2009}.

The observation equation describing the electrophysiological recordings is given by
\begin{equation}\label{eq:ObservationEquation}
	y_t(\mathbf{r}_{n_y}) = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t(\mathbf{r}_{n_y}), 
\end{equation}
where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel at location $\mathbf{r}_{n_y}$ and  $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$  is an i.i.d. Gaussian white noise process and the covariance matrix $\mathbf{\Sigma}_{\epsilon}=\sigma^2_{\epsilon}\mathbf I_{n_y}$, where $\mathbf I$ denotes the identity matrix. 
\section{MRA of the IDE in State-Space}
The multi-resolution approximation \cite{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(r)$ and a mother wavelet $\psi(r)$. Considering a one-dimensional field, the connectivity kernel is decomposed as,
\begin{equation}
 w\left(r-r'\right)=\sum_{l \in \mathbb{Z}}\alpha_{j_0,l}\phi_{j_0,l}\left(r-r'\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l}\psi_{j,l}\left(r-r'\right), 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters, respectively. Scaling functions retain the lowest frequency components, while the wavelet functions extract successively higher frequency components due to their low-pass and band-pass properties, respectively. The field is decomposed using
\begin{equation}
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l}\phi_{j_{0},l}\left(r\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}} \check{x}_{t,j,l}\psi_{j,l}\left(r\right),
\label{eq:FieldExpansion}
\end{equation}
where $x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the coefficients of the expansion and constitute the state vector at time $t$.

@@

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(r-r')=&\left[
    \boldsymbol\phi_{j_0}^\top(r-r') ~
    \boldsymbol\psi_{j_0}^\top(r-r') ~ 
    \boldsymbol\psi_{j_0+1}^\top(r-r') ~
    \cdots ~
    \boldsymbol\psi_{j}^\top(r-r')\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (r) =& \left[
    \boldsymbol\phi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0+1}^\top(r) ~ 
    \cdots ~ 
    \boldsymbol\psi_{j}^\top(r)
\right]   
\end{align}

The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I,
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various scaling and wavelet functions. B-spline functions were chosen due to their compact support and the ability to analytically define the convolution and inner product to form $\boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how $\boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$ are formed, the $m$-th  order cardinal B-spline  function is defined by the recurrence relation \cite{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)dr',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
then a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \cite{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the Cubic spline $\left(m=4\right)$, comprising of third order polynomials added together at the joining points, with an explicit expression given by
\begin{align}
N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
Equation \eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself since
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of \eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$.

The $m$-th order B-spline scaling and wavelet functions are respectively defined as \cite{Chui1992}
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
 where $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients, and
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}.
 \end{align}
By exploiting \eqref{eq:BsplineConvolution} and \eqref{eq:BsplineInnerProduct}, the integrals in \eqref{eq:Lambdax} and \eqref{eq:Lambdatheta} can be computed analytically. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. This can be done for scaling functions by using \eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-k_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 k_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix ...). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using \eqref{eq:MRA-TwoScalepair2}. An example of such an expansion is given in Figure~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.

\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/ScaleDecomposition.pdf}\\
 		\includegraphics[scale=1]{./Graph/WaveletDecomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. The scaling and the wavelet functions (blue) at $j=0$ are respectively obtained using 625 and 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in \eqref{eq:Lambdax} and \eqref{eq:Lambdatheta}.

@@

The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. 
For the linear state-space model given by (\ref{eq:StateEquation}--\ref{eq:MRA-Observationmatrix}), the joint probability distribution $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$ can be written as
 \begin{equation}\label{eq:jointdistribution}
  p(\mathbf X,\mathbf Y;\boldsymbol \Theta)=\prod_{t=0}^{T-1} p(\mathbf y_{t+1}|\mathbf x_{t+1})p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta)p(\mathbf x_0).
 The $\mathcal Q$-function, or lower-bound on the likelihood, can be expressed in terms of the joint distribution components
  \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=\mathbf E_{\boldsymbol \Theta'}\left[2\ln p(\mathbf X,\mathbf Y;\boldsymbol \Theta)\right] \nonumber \\
 &=\mathbf E_{\boldsymbol\Theta'}\left[\sum_{t=0}^{T-1}2\ln p(\mathbf y_{t+1}|\mathbf x_{t+1})+\sum_{t=0}^{T-1}2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta)
 +\sum_{t=0}^{T-1}2\ln p(\mathbf x_0)\right],

>>>>

%    \documentclass[journal,a4paper]{IEEEtran}
 \documentclass[journal]{IEEEtran}

%      \documentclass[10pt,twocolumn,twoside]{IEEEtran}
%     \documentclass[11pt,draftcls,onecolumn,peerreview]{IEEEtran} 
%         \topmargin       -6.0mm
%          \oddsidemargin      0mm
%          \evensidemargin     0mm
%          \textheight     223.5mm
%         \textwidth      170.0mm

@@

\usepackage{stfloats}
\usepackage{float}
\floatstyle{ruled}
\newfloat{algorithm}{htbp}{loa}%[chapter]
\floatname{algorithm}{Algorithm}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Multi-Resolution Neural Field Estimation }

@@

\begin{abstract}
%\boldmath
We develop a multi-resolution approximation (MRA) framework for the IDE neural field equations based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract}
% Note that keywords are not normally used for peerreview papers.
 \begin{IEEEkeywords}
  Neural field model, multiresolution approximation (MRA), Expectation Maximization (EM) algorithm, wavelets.
  \end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
%   \begin{center} \bfseries EDICS Category: SSP-IDEN \end{center}
%   \fi
%
%  For peerreview papers, this IEEEtran command inserts a page break and
%creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

\section{IDE Neural Field Model}
The stochastic IDE form of the Amari neural field  formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an i.i.d. disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$: a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$ \cite{Murphy2009}.
%  due to high computational demands when using a nonlinear (sigmoidal) activation function. }
% \ken{Doing something just for computational efficiency only might be weak ... is there other work we can reference that makes the linearity assumption? ... or are there practical situations when this known to be a good approximations?} 
The observation equation describing the electrophysiological recordings is given by
\begin{equation}\label{eq:ObservationEquation}
	y_t(\mathbf{r}_{n_y}) = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t(\mathbf{r}_{n_y}), 
where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel at location $\mathbf{r}_{n_y}$ and  $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$  is an i.i.d. Gaussian white noise process. % $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel , is corrupted by The superscript $\top$ denotes the transpose operator.
\section{MRA of the IDE in State-Space}
The multi-resolution approximation of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(r)$ and a mother wavelet $\psi(r)$. Considering a one-dimensional field, the connectivity kernel is decomposed as,
 w\left(r-r'\right)=\sum_{l \in \mathbb{Z}}\alpha_{j_0,l}\phi_{j_0,l}\left(r-r'\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l}\psi_{j,l}\left(r-r'\right), 
\label{eq:KernelExpansion}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters, respectively. Scaling functions retain the lowest frequency components, while the wavelet functions extract successively higher frequency components due to their low-pass and band-pass properties, respectively. The field is decomposed using
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l}\phi_{j_{0},l}\left(r\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}} \check{x}_{t,j,l}\psi_{j,l}\left(r\right),
\label{eq:FieldExpansion}
where $x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the coefficients of the expansion and constitute the state vector at time $t$. %\dean{Need to define these guys as the states here!}

@@

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(s)=&\left[
    \boldsymbol\phi_{j_0}^\top(s) ~
    \boldsymbol\psi_{j_0}^\top(s) ~ 
    \boldsymbol\psi_{j_0+1}^\top(s) ~
    \cdots ~
    \boldsymbol\psi_{j}^\top(s)\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (r) =& \left[
    \boldsymbol\phi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0+1}^\top(r) ~ 
    \cdots ~ 
    \boldsymbol\psi_{j}^\top(r)
\right]   
\end{align}

where $s=r-r'$. The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(\mathbf r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I,
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various scaling and wavelet functions. B-spline functions were chosen due to their compact support and the ability to analytically define the convolution and inner product to form $\boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how $\boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$ are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)dr',
\label{SplineConvolutionIntegral}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$.
% \begin{equation}
% N_{1}\left(r\right)=
% \begin{cases}
% 1 & \text{if $ 0\le r<1$}, \\
% 0 & \mathrm{elsewhere}.
% \end{cases}
% \end{equation}
Equation \eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself since
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
A direct consequence of \eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$. The $m^{th}$-order B-spline wavelet is defined as \cite{Chui1992}
 \varphi & _{m}\left(r\right) = \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}}  q_n N_{m}\left(2r-n\right) \\
 q & _n = \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$},
where $q_n$ are the coefficients. By exploiting \eqref{eq:BsplineConvolution} and \eqref{eq:BsplineInnerProduct}, the integrals in \eqref{eq:Lambdax} and \eqref{eq:Lambdatheta} can be computed analytically. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. In this paper, $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in \eqref{eq:Lambdax} and \eqref{eq:Lambdatheta}.

@@

The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. For the linear state-space model given by (\ref{eq:StateEquation}-\ref{eq:Observationmatrix}), the bound used is a quadratic of the form
\begin{equation}\label{eq:Qfunction}
